Radio Emission from Red-Giant Hot Jupiters

### 2015-09-22 ms.tex
cleaned abstract and section 1

\begin{abstract}
%%%
When planet-hosting stars evolve off the main sequence and go through the red-giant branch (RGB), the stars become orders of magnitudes more luminous and at the same time lose mass at much higher rates than their main-sequence counterparts.
Accordingly, planetary companions around them at orbital distances of a few AU, if they exist, will be heated up to the level of canonical hot Jupiters and also subjected to a massive stellar wind.
Given that magnetized planets interacting with stellar winds emit radio waves, such ``Red-Giant Hot Jupiters'' (RGHJs) may also be candidate radio emitters.
We estimate the spectral auroral radio intensity of RGHJs based on the empirical relation with the stellar wind as well as a proposed scaling for planetary magnetic fields. 
We predict that RGHJs might be intrinsically as bright as canonical hot Jupiters, and 100 times brighter than equivalent objects around main-sequence stars, which allow us to search them in 1000 times larger volume. 
The signal from a RGHJ may be detectable at distances up to a few hundred parsecs with current/future instruments such as LOFAR and the Square Kilometer Array.

%Red-giant hot Jupiters (RGHJs) are jovian planets orbiting red-giant-branch (RGB) or asymptotic-giant-branch (AGB) stars. 
%Red-giant hot Jupiters (RGHJs) are jovian planets orbiting red-giant-branch (RGB) or asymptotic-giant-branch (AGB) stars.
%Post-main-sequence stars lose mass at much higher rates than main-sequence stars.
%A jovian planet passing through the dense winds of its RGB or AGB host can capture stellar wind in its magnetosphere, leading to radio emission from electrons spiraling in the magnetic field.
%The cyclotron frequency of electrons from the stellar wind accreting onto the planet scales as 100~MHz~$(B/30 {\rm~Gauss})$.
%The cyclotron frequency of electrons from the stellar wind accreting onto the planet scales as 100~MHz~$(B/30 {\rm~Gauss})$.
%Counterintuitively, a plausible toy model suggests that a stronger planetary magnetic field leads to lower spectral flux density, and might therefore make objects more challenging to observe.
%Still, for a range of planetary and stellar parameters, a RGHJ might generate a radio luminosity that would be visible from kiloparsec distances.
\end{abstract}

@@

Planets with strong magnetic fields may generate radio and/or X-ray emission when interacting with energetic charged particles. 
It is well known that Jupiter emits radio waves from its auroral region due to the cyclotron-maser instability \citep[e.g.][]{wu1979,zarka1998}.  % due to the acceleration of the plasmas. \memoYF{?}.
Potentially, exoplanets can also generate radio emission through similar mechanisms, depending on their intrinsic magnetic fields and the density of surrounding plasmas, e.g. stellar wind particles and particles from Io-like moons. 
Observations of radio emission from Solar System planets imply an empirical relation that the radio intensity is proportional to the input solar wind flux going into each planetary magnetosphere, which is commonly referred to as the ``radiometric Bode's law'' \citep{desch+kaiser1984}. 
Extrapolating this scaling to exoplanets, radio emission from exoplanetary systems has been examined. 
\citet{farrell1999,zarka2001,lazio2004} estimated that a few of the known exoplanetary systems may have radio flux of $\sim $1 mJy level, due to the small orbital distance of the planet and/or the larger planetary mass. 
%\citet{griesmeier2004} modeled in detail the evolution of atmosphere and magnetic field of the tidally-locked close-in stars and found that 
\citet{stevens2005} gave improved estimates of the stellar mass-loss rate based on X-ray flux and re-evaluated the radio flux of known exoplanetary systems. 
\citet{griesmeier2005} took account of the high stellar activity in the early stage of the planetary system and proposed that the young system would be good candidates to search for radio emission. 
\citet{griesmeier2007a, griesmeier2007b} discussed the effects of the detailed properties of stellar wind in the proximity of the stars. They considered not only the kinetic energy of the stellar wind but also the magnetic energy of the stellar wind and coronal mass ejection (CME). 
Note that in these papers, the scaling relation of planetary magnetic fields differ from paper to paper.
In \citet{reiners2010}, the authors adopted a new scaling relation of planetary magnetic field based on \citet{christensen_et_al2009}. 
%\citet{jardine2008} modeled the reconnection between the magnetic fields of close-in planets and of the star, which would accelerate electrons and produce radio emission \memoTM{(can we say?) via cyclotron radiation}.
% I leave this out because this is not based on radiometric bode's law. 
Although observational searches for these radio signatures are underway, no clear detection has been claimed \citep{bastian2000,george2007,stroe2012,hallinan2013,murphy2015}, though there are some promising initial results \citep{lecavelier_et_al2013,sirothia2014}.

@@

In Section \ref{s:assumptions}, we introduce the framework to obtain the frequency of radio emission and the planetary radio emission, and describe our models for the stellar wind and planetary magnetosphere.
Section \ref{s:result} presents an estimate of the spectral radio intensity of RGHJs and compares the predictions with what might be expected from canonical hot Jupiters as well as those from Jupiter-twins.
Section \ref{s:observability} gives the prospects for the signal detection with the current/future instruments. 
Estimates for the actual late-type (M type) evolved stars are also included. 
We find that the Square Kilometer Array(SKA), and perhaps the high band of the LOw Frequency ARray (LOFAR-HBA) as well, will be capable of detecting these signals within a reasonable integration time. 
%In \S5, we discuss the usefulness of radio emission in exploring Jupiter-like planets around evolved stars. 
Section \ref{s:discussion} is devoted to additional discussions, one for the usefulness of the radio emission for characterizing the system and the other for the effects of canceling magnetic field which results from the plasmas spiraling in the planetary magnetosphere.  
Finally, Section \ref{s:conc} concludes the paper with a brief summary.

%Roughly 20\% of the more than 700 \memoYF{check!} currently known exoplanets around main-sequence stars \footnote{See online catalogs such as http://www.openexoplanetcatalogue.com/ \citep{rein2012}, http://exoplanet.eu \citep{schneider_et_al2011}, or http://exoplanets.org \citep{wright_et_al2011} for up-to-date lists.} have masses greater than half of Jupiter's, orbital radii greater than 1~AU, and will become hot Jupiters (i.e., for the present purposes, this means they will receive at least as much irradiation as the currently known hot Jupiters/Neptunes \citep{spiegel+madhusudhan2012}.

@@

Figure \ref{fig:cartoon} is a cartoon for a radio spectrum of a planet with $10~M_J$ and that of the host red giant star with 100~$R_{\odot }$, both of which are placed at a distance of 100 pc. 
Note that the thermal contribution from the accretion disk is negligible in this figure. \memoYF{check} 
The spectral shape is modeled by simply scaling the Jovian radio spectrum (e.g. Figure 8 of \cite{zarka_et_al2004}). 
The continuum lines from the star in the cases of $\alpha_* =0.8$ and $2.0$ are shown as an extreme case. %; the 0.6 value is the prediction from the constant property wind model \citep{wright1975,panagia1975,olnon1975}. 
As Figure \ref{fig:cartoon} indicates, with a reasonable range of spectral index, the radio emission from the star is smaller by order of magnitude than the expected radio flux from RGHJs at frequencies below 300~MHz. 
For less massive planets (with weaker magnetic fields), the distinction is even clearer, as the peak flux is larger and the peak frequencies are smaller.

@@

If we use the assumptions that $R_p \sim R_J\;\;(const.)$, $\rho _c\sim R_p\sim const.$, and $\sigma = const.$, then the assumptions are reduced to:
%%%%%%%%%
\begin{eqnarray}
\mbox{[\#1]} \; P_{\rm rad} &\sim & 4 \times 10^9 \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{0.58} \left( \frac{M_p}{M_J} \right)^{0.98} \left( \frac{a}{a_J} \right)^{-1.17} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{0.586} \left( \frac{v}{v_{\odot}} \right)^{1.46} \\
%
\mbox{[\#2]} \; P_{\rm rad} &\sim & 4 \times 10^{11} \,{\rm W}\, \left( \frac{\omega}{\omega _J} \right)^{0.79} \left( \frac{M_p}{M_J} \right)^{1.33} \left( \frac{a}{a_J} \right)^{-1.60} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{0.8} \left( \frac{v}{v_{\odot}} \right)^{2.0} \\
%
\mbox{[\#3]} \; P_{\rm rad} &\sim & ? \times 10^{11} \,{\rm W}\,\left( \frac{M_p}{M_J} \right)^{2/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
%
\mbox{[\#4]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{2/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#5]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{1/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#6]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{1/2} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#7]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{2/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#8]} \; P_{\rm rad} &\sim & ? \times 10^{11} \,{\rm W}\, \left( \frac{M_p}{M_J} \right)^{0.69}  \left( \frac{t}{4.5 \rm~Gyr} \right)^{-0.288}  \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
\end{eqnarray}
%%%%%%%%%
For the last equation, we used the scaling of the luminosity $L$, equation (\ref{eq:burrowsLum}):
%%%
\begin{equation}
L \sim 1.84\times10^{-10} L_\odot \left( \frac{t}{4.5 \rm~Gyr} \right)^{-1.3} \left( \frac{M_p}{M_J} \right)^{2.64} \tag{\ref{eq:burrowsLum}}
\end{equation} 
%%%
\memoYF{Note that in \citet{ignace2010} the dependence on mass-loss rate and stellar wind velocity is not taken account consistently; in that paper, $P_{\rm rad} \propto \dot M_{\star } v^2$.}

>>>>

\begin{abstract}
%%%
When planet-hosting stars evolve off the main sequence and go through the red-giant branch (RGB), the stars become orders of magnitudes more luminous and at the same time lose mass at much higher rates than their main-sequence counterparts.
Accordingly, planetary companions around them at orbital distances of a few AU, if they exist, will be heated up to the level of canonical hot Jupiters and also subjected to a massive stellar wind.
Given that magnetized planets interacting with stellar winds emit radio waves, such ``Red-Giant Hot Jupiters'' (RGHJs) may also be candidate radio emitters.
We estimate the spectral auroral radio intensity of RGHJs based on the empirical relation with the stellar wind as well as a proposed scaling for planetary magnetic fields. 
We predict that RGHJs might be intrinsically as bright as canonical hot Jupiters, and 100 times brighter than equivalent objects around main-sequence stars, which allow us to search them in 1000 times larger volume. 
The signal from a RGHJ may be detectable at distances up to a few hundred parsecs with current/future instruments such as LOFAR and the Square Kilometer Array. 
\end{abstract}

@@

Planets with strong magnetic fields may generate radio and/or X-ray emission when interacting with energetic charged particles. 
It is well known that Jupiter emits radio waves from its auroral region due to the cyclotron-maser instability \citep[e.g.][]{wu1979,zarka1998}.  Potentially, exoplanets can also generate radio emission through similar mechanisms, depending on their intrinsic magnetic fields and the density of surrounding plasmas, e.g. stellar wind particles and particles from Io-like moons. 
Observations of radio emission from Solar System planets imply an empirical relation that the radio intensity is proportional to the input solar wind flux going into each planetary magnetosphere, which is commonly referred to as the ``radiometric Bode's law'' \citep{desch+kaiser1984}. 
Extrapolating this scaling to exoplanets, radio emission from exoplanetary systems has been examined. 
\citet{farrell1999,zarka2001,lazio2004} estimated that a few of the known exoplanetary systems may have radio flux of $\sim $1 mJy level, due to the small orbital distance of the planet and/or the larger planetary mass. 
\citet{stevens2005} gave improved estimates of the stellar mass-loss rate based on X-ray flux and re-evaluated the radio flux of known exoplanetary systems. 
\citet{griesmeier2005} took account of the high stellar activity in the early stage of the planetary system and proposed that the young system would be good candidates to search for radio emission. 
\citet{griesmeier2007a, griesmeier2007b} discussed the effects of the detailed properties of stellar wind in the proximity of the stars. They considered not only the kinetic energy of the stellar wind but also the magnetic energy of the stellar wind and coronal mass ejection (CME). 
Note that in these papers, the scaling relation of planetary magnetic fields differ from paper to paper.
In \citet{reiners2010}, the authors adopted a new scaling relation of planetary magnetic field based on \citet{christensen_et_al2009}. 
%\citet{jardine2008} modeled the reconnection between the magnetic fields of close-in planets and of the star, which would accelerate electrons and produce radio emission \memoTM{(can we say?) via cyclotron radiation}.
% I leave this out because this is not based on radiometric bode's law. 
Although observational searches for these radio signatures are underway, no clear detection has been claimed \citep{bastian2000,george2007,stroe2012,hallinan2013,murphy2015}, though there are some promising initial results \citep{lecavelier_et_al2013,sirothia2014}.

@@

In Section \ref{s:assumptions}, we introduce the framework to obtain the frequency of radio emission and the planetary radio emission, and describe our models for the stellar wind and planetary magnetosphere.
Section \ref{s:result} presents an estimate of the spectral radio intensity of RGHJs and compares the predictions with what might be expected from canonical hot Jupiters as well as those from Jupiter-twins.
Section \ref{s:observability} gives the prospects for the signal detection with the current/future instruments. 
Estimates for the actual late-type (M type) evolved stars are also included. 
We find that the Square Kilometer Array(SKA), and perhaps the high band of the LOw Frequency ARray (LOFAR-HBA) as well, will be capable of detecting these signals within a reasonable integration time. 
Section \ref{s:discussion} is devoted to additional discussions, one for the usefulness of the radio emission to characterize the system and the other for the effects of canceling magnetic field which results from the plasmas spiraling in the planetary magnetosphere.  
Finally, Section \ref{s:conc} concludes the paper with a brief summary.

@@

Figure \ref{fig:cartoon} is a cartoon for a radio spectrum of a planet with $10~M_J$ and that of the host red giant star with 100~$R_{\odot }$, both of which are placed at a distance of 100 pc. 
Note that the thermal contribution from the accretion disk is negligible in this figure. 
The spectral shape is modeled by simply scaling the Jovian radio spectrum (e.g. Figure 8 of \cite{zarka_et_al2004}). 
The continuum lines from the star in the cases of $\alpha_* =0.8$ and $2.0$ are shown as an extreme case. %; the 0.6 value is the prediction from the constant property wind model \citep{wright1975,panagia1975,olnon1975}. 
As Figure \ref{fig:cartoon} indicates, with a reasonable range of spectral index, the radio emission from the star is smaller by order of magnitude than the expected radio flux from RGHJs at frequencies below 300~MHz. 
For less massive planets (with weaker magnetic fields), the distinction is even clearer, as the peak flux is larger and the peak frequencies are smaller.

@@

If we use the assumptions that $R_p \sim R_J\;\;(const.)$, $\rho _c\sim R_p\sim const.$, and $\sigma = const.$, then the assumptions are reduced to:
%%%%%%%%%
\begin{eqnarray}
\mbox{[\#1]} \; P_{\rm rad} &\sim & 4 \times 10^9 \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{0.58} \left( \frac{M_p}{M_J} \right)^{0.98} \left( \frac{a}{a_J} \right)^{-1.17} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{0.586} \left( \frac{v}{v_{\odot}} \right)^{1.46} \\
%
\mbox{[\#2]} \; P_{\rm rad} &\sim & 4 \times 10^{11} \,{\rm W}\, \left( \frac{\omega}{\omega _J} \right)^{0.79} \left( \frac{M_p}{M_J} \right)^{1.33} \left( \frac{a}{a_J} \right)^{-1.60} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{0.8} \left( \frac{v}{v_{\odot}} \right)^{2.0} \\
%
\mbox{[\#3]} \; P_{\rm rad} &\sim & ? \times 10^{11} \,{\rm W}\,\left( \frac{M_p}{M_J} \right)^{2/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
%
\mbox{[\#4]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{2/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#5]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{1/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#6]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{1/2} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#7]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{2/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#8]} \; P_{\rm rad} &\sim & ? \times 10^{11} \,{\rm W}\, \left( \frac{M_p}{M_J} \right)^{0.69}  \left( \frac{t}{4.5 \rm~Gyr} \right)^{-0.288}  \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
\end{eqnarray}
%%%%%%%%%
For the last equation, we used the scaling of the luminosity $L$, equation (\ref{eq:burrowsLum}):
%%%
\begin{equation}
L \sim 1.84\times10^{-10} L_\odot \left( \frac{t}{4.5 \rm~Gyr} \right)^{-1.3} \left( \frac{M_p}{M_J} \right)^{2.64} \tag{\ref{eq:burrowsLum}}
\end{equation} 
%%%
